Summer 2024 Report

### 2024-09-12 Relations/Report.tex
ARS

\newcommand{\tto}{\twoheadrightarrow}

@@

\subsection{Well-foundedness}
\textcolor{red}{Talk about the different types of wellfoundedness, starting with the one given in the book, the reason for thinking it is classical and not constructive, and provide an image showing the relation between the notions of wellfoundedness we have. Talk about future work in this area.}
\\
To date we have investigated the differences between four distinct notions of well-foundedness. The make clear the difference between these distinct notions of well-foundedness we first present the following defintiions for a given relation $R$.
\begin{enumerate}
    \item \textit{Accessible:} An element $x$ is \textit{R-accessible} if all elements that have an $R$ reduction to $x$ are themselves \textit{R-accessible}.  
    \item \textit{Inductive predicate:} A predicate $\varphi$ is an inductive predicate if $\varphi \, x$ is true whenever $\varphi$ is true for all elements that have an $R$ reduction to $x$.
    \item \textit{Minimal element:} An element $x$ is minimal for a given predicate $\varphi$ if $\varphi \, x$ is true and no other element for which $\varphi$ holds has an $R$ reduction to $x$. 
    \item \textit{Decreasing:} A sequence $s$ is decreasing if every $R$ reduction decreases the size of $n$ (where $n \in \mathbb{N}$).   \textcolor{red}{bad defn. Improve.}
\end{enumerate}
With the above definitions established we define the notions of well-foundedness we have been investigating. For a given relation $R$:
\begin{enumerate}
    \item \textit{R} is well-founded if every element is \textit{R}-accessible.
    \item \textit{R} is well-founded if every inductive predicate is universally true.
    \item \textit{R} is well-founded if every non-empty predicate has a minimal element with respect to $R$.
    \item \textit{R} is well-founded if every sequence contains a non-decreasing index.
\end{enumerate}
Further, we explored weaker variants of these notions of well-foundedness:
\begin{enumerate}
    \item \textit{R} is well-founded if every element is $\lnot \lnot$  \textit{R}-accessible.
    \item \textit{R} is well-founded if for every inductive predicate $\varphi$, $\lnot \lnot \varphi$ is universally true.
    \item \textit{R} is well-founded if it is not not the case that every non-empty predicate has a minimal element with respect to $R$. \textcolor{red}{is there a better way to express this double negation in english?} 
    \item \textit{R} is well-founded if there is no infinite $R$-decreasing sequence (this is the definition of well-foundedness found in TeReSe).
\end{enumerate}

>>>>

\newcommand{\tto}{\twoheadrightarrow}
\newcommand{\sse}{\subseteq}
\newcommand{\bset}{\mathbf{Set}}

\newcommand{\sacomment}[1]{\textcolor{red}{#1}}
\newcommand{\apcomment}[1]{\textcolor{blue}{#1}}

\usepackage{amsthm}

\newtheorem{thm}{Theorem}
\newtheorem{dfn}[thm]{Definition}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lemma}[thm]{Lemma}
\newtheorem{rmk}[thm]{Remark}
\newtheorem{expl}[thm]{Example}
\newtheorem{notn}[thm]{Notation}
%\theoremstyle{nonumberplain}
%\theoremsymbol{\Box}
% \newtheorem{proof}{Proof}

@@

\subsection{Well-foundedness}
\sacomment{Talk about the different types of wellfoundedness, starting with the one given in the book, the reason for thinking it is classical and not constructive, and provide an image showing the relation between the notions of wellfoundedness we have. Talk about future work in this area.}

\apcomment{Mention that, for a set $A$, we will use "subset of $A$" and "predicate on $A$" interchangeably,
with subsets being represented in the Agda code using predicates, which are in turn formalized as
dependent types over $A$, i.e. maps $\varphi : A \to \bset$ }

Working in constructive logic introduces subtle differences between notions of well-foundedness of relations
which are not present in classical logic.  In our investigation, we have focused on four distinct definitions of well-foundedness, with minor variations of each. To spell these out, we first define some preliminary notions.
% The make clear the difference between these distinct notions of well-foundedness we first present the following defintiions for a given relation $R$.

\begin{dfn} Let $A$ be a set and $R \sse A \times A$ be a relation.
\begin{enumerate}
  \item Let $\phi \sse A$.  An element $x$ is $R$-minimal for $\varphi$ if $\varphi(x)$ is true and no other element for which $\varphi$ holds is $R$-related to $x$.
  \item An element $x \in A$ is \textit{$R$-accessible} if every element $R$-related to $x$ is \textit{$R$-accessible}.  (This should be understood as an inductive definition, with base case consisting of $R$-minimal elements for the whole of $A$.)
    \item A predicate $\varphi$ on $A$ is \emph{inductive} if it holds on a given element $x \in A$ whenever it holds on all elements $R$-related to $x$:
    \[ (\forall y. Ryx \to \phi(y)) \to \phi(x) \]
    \item A sequence $s = (s_0,s_1,\dots)$ of elements of $A$ is $R$-decreasing
    if every element is related to its predecessor.  That is, $Rs_{n+1}s_n$ for all $n$.
\end{enumerate}
\end{dfn}
With the above definitions established we define the notions of well-foundedness we have been investigating. For a given relation $R$:
\begin{enumerate}
    \item \textit{R} is well-founded if every element is \textit{R}-accessible.
    \item \textit{R} is well-founded if every inductive predicate is universally true.
    \item \textit{R} is well-founded if every non-empty predicate has a minimal element with respect to $R$.
    \item \textit{R} is well-founded if every sequence contains a non-decreasing index.
\end{enumerate}
Further, we explored weaker variants of these notions of well-foundedness:
\begin{enumerate}
    \item \textit{R} is well-founded if every element is $\lnot \lnot$  \textit{R}-accessible.
    \item \textit{R} is well-founded if for every inductive predicate $\varphi$, $\lnot \lnot \varphi$ is universally true.
    \item \textit{R} is well-founded if it is not not the case that every non-empty predicate has a minimal element with respect to $R$. \textcolor{red}{is there a better way to express this double negation in english?}
    \item \textit{R} is well-founded if there is no infinite $R$-decreasing sequence (this is the definition of well-foundedness found in TeReSe).
\end{enumerate}
